---
title: "Peer_graded_Assignment2"
author: "Toru Hasegawa"
date: "2019/6/1"
output: html_document
---
#The Worst Weather Event for Population Health and Economic

##Synopsis
In this assignment, I analyze  [Storm Data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2) based on the [National Weather Service Storm Data](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf).

And the subject of this analysis is clearing up following point,
1. What type of event is the worst for population health
2. What type of event is the worst for economic


##Data Processing
```{r download_data, echo=TRUE}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2",destfile = "Stome_data.bz2")
```

```{r extract_and_setdata, echo=TRUE}
storm_data <- read.csv("Stome_data.csv.bz2")

head(storm_data)
```


```{r prepare_data, echo=TRUE}
columnlist <- c("EVTYPE", "FATALITIES", "INJURIES","PROPDMG","PROPDMGEXP", "CROPDMG","CROPDMGEXP")
df <- storm_data[,columnlist]

summary(df)
```

```{r check_data, echo=TRUE}
table(is.na(df))
#The dataset don't have NA data.
```


##Results

1. Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?

```{r barplot_health, echo=TRUE}
FatalitiesPerEvtype <- aggregate(FATALITIES~EVTYPE, df, FUN=sum)
InjuriesPerEvtype <-aggregate(INJURIES~EVTYPE, df, FUN=sum)

#Search the most harmful event for health
FatalitiesPerEvtype$EVTYPE[which.max(FatalitiesPerEvtype$FATALITIES)]
InjuriesPerEvtype$EVTYPE[which.max(InjuriesPerEvtype$INJURIES)]

#make barplot of top6 harmful event
sort <- order(FatalitiesPerEvtype$FATALITIES, decreasing = TRUE)
Ordered_Fatalities <- FatalitiesPerEvtype[sort,]
sort2 <- order(InjuriesPerEvtype$INJURIES, decreasing = TRUE)
Ordered_Injuries <- InjuriesPerEvtype[sort2,]

par(mfrow=c(2,1))
barplot(head(Ordered_Fatalities$FATALITIES), names.arg = head(Ordered_Fatalities$EVTYPE), cex.names = 0.5, main="Event of Fatalities")
barplot(head(Ordered_Injuries$INJURIES), names.arg = head(Ordered_Injuries$EVTYPE), cex.names = 0.5, main="Event of Injuries")
```

As result, the top of harmful event to population health is "TORNADO"" in both of fatalities and injuries aspect.


2. Across the United States, which types of events have the greatest economic consequences?
```{r check_levels, echo=TRUE}
summary(df$PROPDMGEXP)
summary(df$CROPDMGEXP)
```

Based on the above results, this data have a lot of metric prefix.
Exchange those metric prerfix to number
```{r exchange_metric_perfix, echo=TRUE}
#exchange crop data
df$CROPDMGEXP <- as.character(df$CROPDMGEXP)
df$CROPDMGEXP[df$CROPDMGEXP == "0" | df$CROPDMGEXP == "2" ] <- "10"
df$CROPDMGEXP[df$CROPDMGEXP == "?" | df$CROPDMGEXP == "" ] <- "0"
df$CROPDMGEXP[df$CROPDMGEXP == "B" ] <- "1000000000"
df$CROPDMGEXP[df$CROPDMGEXP == "m" | df$CROPDMGEXP == "M" ] <- "1000000"
df$CROPDMGEXP[df$CROPDMGEXP == "k" | df$CROPDMGEXP == "K" ] <- "1000"
df$CROPDMGEXP <- as.numeric(df$CROPDMGEXP)
df$CROPDMG2 <- df$CROPDMGEXP * df$CROPDMG

#exchange prop data
df$PROPDMGEXP <- as.character(df$PROPDMGEXP)
df$PROPDMGEXP[df$PROPDMGEXP == "0" | df$PROPDMGEXP == "1" ] <- "10"
df$PROPDMGEXP[df$PROPDMGEXP == "2" | df$PROPDMGEXP == "3" ] <- "10"
df$PROPDMGEXP[df$PROPDMGEXP == "4" | df$PROPDMGEXP == "5" ] <- "10"
df$PROPDMGEXP[df$PROPDMGEXP == "6" | df$PROPDMGEXP == "7" ] <- "10"
df$PROPDMGEXP[df$PROPDMGEXP == "8" ] <- "10"
df$PROPDMGEXP[df$PROPDMGEXP == "?" | df$PROPDMGEXP == "" ] <- "0"
df$PROPDMGEXP[df$PROPDMGEXP == "-" ] <- "0"
df$PROPDMGEXP[df$PROPDMGEXP == "+" ] <- "1"
df$PROPDMGEXP[df$PROPDMGEXP == "B" ] <- "1000000000"
df$PROPDMGEXP[df$PROPDMGEXP == "m" | df$PROPDMGEXP == "M" ] <- "1000000"
df$PROPDMGEXP[df$PROPDMGEXP == "k" | df$PROPDMGEXP == "K" ] <- "1000"
df$PROPDMGEXP[df$PROPDMGEXP == "h" | df$PROPDMGEXP == "H" ] <- "100"
df$PROPDMGEXP <- as.numeric(df$PROPDMGEXP)
df$PROPDMG2 <- df$PROPDMGEXP * df$PROPDMG
```


```{r worst_event_for_prop, echo=TRUE}
#prepare dataset to make ranking
prop_sum <- aggregate(PROPDMG2~EVTYPE, df,FUN=sum)

#Search the worst event for prop
sort3 <- order(prop_sum$PROPDMG2, decreasing = TRUE)
Ordered_prop_sum <- prop_sum[sort3,]
head(Ordered_prop_sum)
```

As result, it is obvious that the worst event type for property is "FLOOD".
The damage of economic is 144.5 billion $.


```{r worst_event_for_crop, echo=TRUE}
#prepare dataset to make ranking
crop_sum <- aggregate(CROPDMG2~EVTYPE, df, FUN=sum)

#Search the worst event for crop
sort4 <- order(crop_sum$CROPDMG2, decreasing = TRUE)
Ordered_crop_sum <- crop_sum[sort4,]
head(Ordered_crop_sum)
```

The top of worst event for crop is "DROUGHT".

Sum up of those results, and calcurate total economic effect for each event types.
```{r barplot_total_effect_for_ecconomic, echo=TRUE}
#merge prop and crop data
total_effect <- merge(prop_sum, crop_sum)
total_effect$total_sum <- total_effect$PROPDMG2 + total_effect$CROPDMG2
sort5<- order(total_effect$total_sum, decreasing = TRUE)
Ordered_total_effect <-total_effect[sort5,]

#make barplot
barplot(head(Ordered_total_effect$total_sum), names.arg = head(Ordered_total_effect$EVTYPE), cex.names = 0.5, main="Top harmful event for economic")
```
In conclusion, "FLOOD"is most harmful event for economic.

